\bf Model Visualization Techniques for a Social Network Model

### 2017-03-27 paper.Rnw
some bg on net viz

\par The guiding principles of model visualization from \citet{modelvis} are (1) display the model in the data space, (2) collections of models are more informative than singletons, and (3) explore how algorithms work, instead of just looking at the final result. Stochastic actor-oriented models are a prime example of a set of models that  benefit greatly from application of model visualization. First, the models themselves include a continuous-time Markov chain (CTMC) that is completely hidden in the model fitting process. Bringing the CTMC out from behind-the-scenes and into the light of visualization  provides researchers with insights into the underlying features of the model. Secondly, SAOMs allow for a great deal of parameters that can be added to the objective function, each of which are attached to a network statistic. These statistics are often somewhat, if not highly, correlated, and so are the associated parameters. By visualizing collections of SAOMs, we can gain a better understanding of these correlations and find ways to deal with them and rectify their effects. Finally, estimation of the parameters in a SAOM relies on a Robbins-Monro algorithm, and convergence checks rely on simulations from the fitted model. Again, in these steps are largely hidden from the researcher computing them with software. We hope that by bringing these background elements and others into the foregroundwith static and dynamic visualizations, stochastic actor-oriented models will be better understood and more accessible to social network analysts.

>>>>

\par The guiding principles of model visualization from \citet{modelvis} are (1) display the model in the data space, (2) collections of models are more informative than singletons, and (3) explore how algorithms work, instead of just looking at the final result. Stochastic actor-oriented models are a prime example of a set of models that  benefit greatly from application of model visualization. First, the models themselves include a continuous-time Markov chain (CTMC) that is completely hidden in the model fitting process. Bringing the CTMC out from behind-the-scenes and into the light of visualization  provides researchers with insights into the underlying features of the model. Secondly, SAOMs allow for a great deal of parameters that can be added to the objective function, each of which are attached to a network statistic. These statistics are often somewhat, if not highly, correlated, and so are the associated parameters. By visualizing collections of SAOMs, we can gain a better understanding of these correlations and find ways to deal with them and rectify their effects. Finally, estimation of the parameters in a SAOM relies on a Robbins-Monro algorithm, and convergence checks rely on simulations from the fitted model. Again, in these steps are largely hidden from the researcher computing them with software. We hope that by bringing these background elements and others into the foreground with static and dynamic visualizations, stochastic actor-oriented models will be better understood and more accessible to social network analysts.

\section{Methods for Network Visualization}

Network visualization, also called network mapping, is a prominent subfield of network analysis. Visualizing network data is uniquely difficult because of the structure of the data itself.  Most visualizations rely on a well-defined axis inherited from the data. If data are numerical, histograms, scatterplots, or time series plots are straightforward to construct. If the data are categorical, bar charts and mosaic plots can be constructed easily. If the data are spatial, there is a well-defined space in which to view information. Network data are much less cut-and-dried.

There are two primary methods used to visualize networks. The first method represents nodes with points in 2D Euclidean space and then represents edges by connecting the points with lines when there is an edge between the two nodes. But, because there is no natural placement of the points, a random placement is used, then adjusted iteratively via a layout algorithm, of which there are many.

Some layout algorithms were designed to mimic physical systems, drawing the graphs based on the "forces" connecting them. The network's edges act as springs pushing and pulling the nodes in 2D space. Others use properties of the network's adjaceny matrix in order to place the nodes. Some commonly used layout algorithms are  XX ADD MORE DETAIL LATER XX
\begin{itemize}
\item Kamada-Kawai
\item Fructerman-Reingold
\item Multi-dimensional scaling
\item Eigen
\end{itemize}

The second primary method for network visualization uses the adjacency matrix of the network. This method is particularly useful when the network is very complex, dense, or large. XX MORE STUFF ABOUT ADJ MAT VIZ XX
